---
title: "Latency analysis for Barrier"
author: "Giulio Fantuzzi"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
```


# Barrier analysis

First of all, import the datasets i obtained from *ORFEO*
```{r}
barrier_default.df<- read.csv(file = "barrier_default.csv")
barrier_linear.df<- read.csv(file = "barrier_linear.csv")
barrier_doublering.df<- read.csv(file = "barrier_doublering.csv")
barrier_recursivedoubling.df<- read.csv(file = "barrier_recursivedoubling.csv")
barrier_bruck.df<- read.csv(file = "barrier_bruck.csv")
barrier_tree.df<- read.csv(file = "barrier_tree.csv")
```

Now we can merge them into a single df:
```{r}
barrier.df<- rbind(barrier_default.df,barrier_linear.df,barrier_doublering.df,barrier_recursivedoubling.df,barrier_bruck.df,barrier_tree.df)
```


# Fix algorithm and see allocation effect

### Linear Barrier

```{r,fig.width=8, fig.height=5,echo=F}
barrier.df %>% filter(Algorithm =="Linear")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Linear Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```

### Tree Barrier

```{r,fig.width=8, fig.height=5,echo=F}
barrier.df %>% filter(Algorithm =="Tree")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Tree Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```

### Bruck Barrier

```{r,fig.width=8, fig.height=5,echo=F}
barrier.df %>% filter(Algorithm =="Bruck")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Bruck Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```

### Default Barrier

```{r,fig.width=8, fig.height=5,echo=F}
barrier.df %>% filter(Algorithm =="Default")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Default Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```


# Fixing allocation and comparing algorithms

```{r,fig.width=15,echo=F}
Plots=list()
maps= c("core","socket","node")
for (i in 1:length(maps)){
  Plots[[i]]<- barrier.df %>% filter(Allocation == maps[i]) %>% filter(Algorithm %in% c("Default","Tree","Linear","Bruck")) %>%
    ggplot(aes(x = Processes, y = Latency, color = Algorithm)) +
      geom_line(linetype = 1) +
      geom_point(size=1)+
      geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
      annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
      scale_color_manual(values = c("#E67E22","#8E44AD","#27AE60","#3498DB")) +
      xlim(2, 48) +  # Adjust the x-axis limits as needed
      labs(x = 'N. of processes (#cores)',
           y = 'Latency(μs)',
           title = paste0('Barrier Latency vs #cores'),
           color = 'Algorithm')+
      theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
}
Plots[[1]] | Plots[[2]] | Plots[[3]]
```


# Latency models

## Hockney model for linear barrier

```{r}
#Point to point latency (alpha parameter in Hockney model for linear barrier)
alpha<- 0.19
# Model will be Latency= alpha*(P-1) = alpha*P - alpha, with P: number of processes
barrier_linear.df$HockneyLatency <- alpha* barrier_linear.df$Processes - alpha
barrier_linear.df$ObservedLatency <-barrier_linear.df$Latency
# Now make the plot
Hockney.df <- tidyr::gather(barrier_linear.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(Hockney.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(Hockney.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(Hockney.df, Legend == "HockneyLatency"), linetype = 1) +
  labs(title = expression(paste("Linear Barrier: Hockney model fit (with ",hat(alpha),"= 0.19 μs)")),
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey19", "HockneyLatency" = "#00BFC1")) +
  theme_minimal()
```


NB: here we don't need to log2() the latency...the skewness of the variable was given by the effect of different message sizes, but here we don't consider the size anymore!!!

## Linear models for all the algorithms

### Linear Barrier Model
```{r}
linear_fit<- lm(formula= Latency ~ -1+Processes,
                data=barrier_linear.df %>% filter(Allocation == "core")
                )
summary(linear_fit)
```

```{r,echo=F}
barrier_linear.df$PredictedLatency <- predict(linear_fit)
barrier_linear.df$ObservedLatency <-barrier_linear.df$Latency
# Now make the plot
LinearPlot.df <- tidyr::gather(barrier_linear.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(LinearPlot.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(LinearPlot.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(LinearPlot.df, Legend == "PredictedLatency"), linetype = 1) +
  labs(title = "Linear barrier: estimated model fit",
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey30", "PredictedLatency" = "#27AE60")) +
  theme_minimal()+
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11) )
```

For the non linear algorithms I noticed that introducing a quadratic term increase the fit:

```{r}
tree_fit<- lm(formula=Latency ~ -1+Processes+I(Processes^2),
                data=barrier_tree.df %>% filter(Allocation == "core")
                )
summary(tree_fit)
```


```{r,echo=F}
barrier_tree.df$PredictedLatency <- predict(tree_fit)
barrier_tree.df$ObservedLatency <-barrier_tree.df$Latency
# Now make the plot
TreePlot.df <- tidyr::gather(barrier_tree.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(TreePlot.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(TreePlot.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(TreePlot.df, Legend == "PredictedLatency"), linetype = 1) +
  labs(title = "Tree barrier: estimated model fit",
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey30", "PredictedLatency" = "#3498DB"),) +
  theme_minimal()+
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11) )
```

### Bruck Barrier Model
```{r}
bruck_fit<- lm(formula= Latency ~ -1+Processes+I(Processes^2),
                data=barrier_bruck.df %>% filter(Allocation == "core")
                )
summary(bruck_fit)
```


```{r,echo=F}
barrier_bruck.df$PredictedLatency <- predict(bruck_fit)
barrier_bruck.df$ObservedLatency <-barrier_bruck.df$Latency
# Now make the plot
BruckPlot.df <- tidyr::gather(barrier_bruck.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(BruckPlot.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(BruckPlot.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(BruckPlot.df, Legend == "PredictedLatency"), linetype = 1) +
  labs(title = "Bruck barrier: estimated model fit",
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey30", "PredictedLatency" = "#E67E22")) +
  theme_minimal()+
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11) )
```

### Default Barrier Model

```{r}
default_fit<- lm(formula= Latency ~ -1+Processes+I(Processes^2),
                data=barrier_default.df %>% filter(Allocation == "core")
                )
summary(default_fit)
```


```{r,echo=F}
barrier_default.df$PredictedLatency <- predict(default_fit)
barrier_default.df$ObservedLatency <-barrier_default.df$Latency
# Now make the plot
DefaultPlot.df <- tidyr::gather(barrier_default.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(DefaultPlot.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(DefaultPlot.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(DefaultPlot.df, Legend == "PredictedLatency"), linetype = 1) +
  labs(title ="Default barrier: estimated model fit",
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey30", "PredictedLatency" = "#8E44AD")) +
  theme_minimal()+
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11) )
```