---
title: "Barrier Analysis"
author: "Giulio Fantuzzi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
```


# Barrier analysis

First of all, import the datasets i obtained from *ORFEO*
```{r}
barrier_default.df<- read.csv(file = "barrier_default.csv")
barrier_linear.df<- read.csv(file = "barrier_linear.csv")
barrier_doublering.df<- read.csv(file = "barrier_doublering.csv")
barrier_recursivedoubling.df<- read.csv(file = "barrier_recursivedoubling.csv")
barrier_bruck.df<- read.csv(file = "barrier_bruck.csv")
barrier_tree.df<- read.csv(file = "barrier_tree.csv")
```

Now we can bind them into a single df:
```{r}
barrier.df<- rbind(barrier_default.df,barrier_linear.df,barrier_doublering.df,barrier_recursivedoubling.df,barrier_bruck.df,barrier_tree.df)
```


# Fix algorithm and see allocation effect
```{r,fig.width=8, fig.height=5}
barrier.df %>% filter(Algorithm =="Linear")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Linear Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```

COMMENTS: as we saw for broadcast, the 3 lines seem to converge (the more we fill the cores, the less the allocation will have an impact). Another interesting thing is that here we don't have any jump. I suppose this is because here with barrier we are communicating a 0 sized message...


```{r,fig.width=8, fig.height=5}
barrier.df %>% filter(Algorithm =="Tree")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Tree Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```


```{r,fig.width=8, fig.height=5}
barrier.df %>% filter(Algorithm =="Bruck")%>% 
  ggplot(aes(x = Processes, y = Latency, color = Allocation)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    scale_color_manual(values = c("#27AE60","#E67E22","#3498DB")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Bruck Barrier Latency vs #cores (by Allocation)',
         color = 'Allocation')+
    theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
```


# Fix allocation and see algorithms comparison

```{r,fig.width=8, fig.height=5}
barrier.df %>% filter(Allocation == "core") %>% filter(Algorithm %in% c("Tree","Linear","Bruck") )%>% 
  ggplot(aes(x = Processes, y = Latency, color = Algorithm)) +
    geom_line(linetype = 1) +
    geom_point(size=1)+
    #geom_abline(slope = 0.19, intercept = -0.19, color = "red")+
    geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
    annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
    #scale_color_manual(values = c("#27AE60","#E67E22","#3498DB","#8E44AD","")) +
    xlim(2, 48) +
    labs(x = 'N. of processes (#cores)',
         y = 'Latency(μs)',
         title = 'Barrier Latency vs #cores (by Algorithm)',
         color = 'Algorithm')+
    theme_light()
```

# Fixing allocation and comparing algorithms

```{r}
Plots=list()
maps= c("core","socket","node")
for (i in 1:length(maps)){
  Plots[[i]]<- barrier.df %>% filter(Allocation == maps[i]) %>% filter(Algorithm %in% c("Default","Tree","Linear","Bruck")) %>%
    ggplot(aes(x = Processes, y = Latency, color = Algorithm)) +
      geom_line(linetype = 1) +
      geom_point(size=1)+
      geom_vline(xintercept = c(12, 24), linetype = "dashed", color = "#616A6B")+
      annotate("text", x = c(12,24), y = c(0,0), label = c("12 Cores","24 cores"), vjust = 0, hjust = -0.1,color="#616A6B")+
      scale_color_manual(values = c("#E67E22","#8E44AD","#27AE60","#3498DB")) +
      xlim(2, 48) +  # Adjust the x-axis limits as needed
      labs(x = 'N. of processes (#cores)',
           y = 'Latency(μs)',
           title = paste0('Barrier Latency vs #cores'),
           color = 'Algorithm')+
      theme_light()+
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 12) 
    )
}
```

```{r,fig.width=15}
Plots[[1]] | Plots[[2]] | Plots[[3]]
```


# Latency models

First of all let's present the hockney model for the linear barrier algorithm:
```{r}
#Point to point latency (alpha parameter in Hockney model for linear barrier)
alpha<- 0.19
# Model will be Latency= alpha*(P-1) = alpha*P - alpha, with P: number of processes
barrier_linear.df$HockneyLatency <- alpha* barrier_linear.df$Processes - alpha
barrier_linear.df$ObservedLatency <-barrier_linear.df$Latency
# Now make the plot
Hockney.df <- tidyr::gather(barrier_linear.df %>% filter(Allocation=="core"), key = "Legend", value = "Value", -Processes, -Allocation, -Algorithm,-Latency)
# Plot using ggplot
ggplot(Hockney.df, aes(x = Processes, y = Value, color = Legend)) +
  geom_point(data = subset(Hockney.df, Legend == "ObservedLatency"), size = 1) +
  geom_line(data = subset(Hockney.df, Legend == "HockneyLatency"), linetype = 1) +
  labs(title = expression(paste("Linear Barrier: Hockney model fit (with ",hat(alpha),"= 0.19 μs)")),
       x = "N. of Processes",
       y = "Lacency(μs)") +
  scale_color_manual(values = c("ObservedLatency" = "grey19", "HockneyLatency" = "#00BFC1")) +
  theme_minimal()
```


NB: here we don't need to log2() the latency...the skewness of the variable was given by the effect of different message sizes, but here we don't consider the size anymore!!!

```{r}
linear_fit<- lm(formula= Latency ~ -1+Processes,
                data=barrier_linear.df %>% filter(Allocation == "core")
                )
summary(linear_fit)
```

For the non linear algorithms I noticed that introducing a quadratic term increase the fit:

```{r}
tree_fit<- lm(formula=Latency ~ -1+Processes+I(Processes^2),
                data=barrier_tree.df %>% filter(Allocation == "core")
                )
summary(tree_fit)
```
```{r}
bruck_fit<- lm(formula= Latency ~ -1+Processes+I(Processes^2),
                data=barrier_bruck.df %>% filter(Allocation == "core")
                )
summary(bruck_fit)
```


```{r}
default_fit<- lm(formula= Latency ~ -1+Processes+I(Processes^2),
                data=barrier_default.df %>% filter(Allocation == "core")
                )
summary(default_fit)
```